### 2015-11-30 manuscript/tables/ideology_regressions.tex
First cut at ideology analysis
- see updated datasets on HPC

% Table created by stargazer v.5.2 by Marek Hlavac, Harvard University. E-mail: hlavac at fas.harvard.edu
% Date and time: Mon, Nov 30, 2015 - 11:50:41
\begin{table}[!htbp] \centering 
  \caption{Regression results for log-linear models for ideology
analysis. The models from left to right: 1) Alignment score
by section (dependence not considered yet),
2) Sum of alignment scores of all sections of
bill dyads, 3) Number of alignments of bill dyad, 4)
mean alignment score of dyad, 5) penalized mean alignment
score of dyad, 5) maximum alignment score of dyad. } 
  \label{tab:ideo_reg} 
\begin{tabular}{@{\extracolsep{5pt}}lcccccc} 
\\[-1.8ex]\hline 
\hline \\[-1.8ex] 
 & \multicolumn{6}{c}{\textit{Dependent variable:}} \\ 
\cline{2-7} 
\\[-1.8ex] & Score & Sum & Number & Mean & Penal & Max \\ 
\\[-1.8ex] & (1) & (2) & (3) & (4) & (5) & (6)\\ 
\hline \\[-1.8ex] 
 ideology\_dist & $-$0.029$^{***}$ & $-$0.034$^{***}$ & $-$0.009$^{***}$ & $-$0.025$^{***}$ & 0.002 & $-$0.031$^{***}$ \\ 
  & (0.0003) & (0.0005) & (0.0003) & (0.0004) & (0.002) & (0.0004) \\ 
  & & & & & & \\ 
 Constant & 3.745$^{***}$ & 4.300$^{***}$ & 0.486$^{***}$ & 3.814$^{***}$ & 2.982$^{***}$ & 3.972$^{***}$ \\ 
  & (0.001) & (0.001) & (0.001) & (0.001) & (0.004) & (0.001) \\ 
  & & & & & & \\ 
\hline \\[-1.8ex] 
Observations & 3,957,875 & 1,745,336 & 1,745,336 & 1,745,336 & 1,745,336 & 1,745,336 \\ 
Adjusted R$^{2}$ & 0.003 & 0.003 & 0.001 & 0.003 & 0.00000 & 0.003 \\ 
\hline 
\hline \\[-1.8ex] 
\textit{Note:}  & \multicolumn{6}{r}{$^{*}$p$<$0.1; $^{**}$p$<$0.05; $^{***}$p$<$0.01} \\ 
\end{tabular} 
\end{table}