### 2016-01-12 manuscript/tables/ideology_regs.tex
more on substantive interprtation of regression results

% latex table generated in R 3.2.2 by xtable 1.7-4 package
% Mon Dec 21 04:04:01 2015
\begin{table}[ht]
\centering
\begin{tabular}{rrrrr}
  \hline
 & Intercept & Estimate & Std.Dev & p.value \\ 
  \hline
Sum & 4.300 & -0.034 & 0.001 & 0.000 \\ 
  Mean & 3.814 & -0.025 & 0.001 & 0.000 \\ 
  Max & 3.972 & -0.031 & 0.001 & 0.000 \\ 
  No. Alignments & 0.486 & -0.009 & 0.001 & 0.000 \\ 
   \hline
\end{tabular}
\caption{Log-Linear models for alignment and euclidian
distance in ideology. Two tailed p-values are generated from quadratic
assignment procedure with 1000 iterations. Std.Dev is the standard
deviation of the null distribution. The rows correspond to
different methods of aggregating the section alignments to bill
allignments: \textit{Sum}: Sum of alignment scores of all alignments,
\textit{Mean}: Average alignment score accross secion pairs bill dyad,
\textit{Max}: Highest alignment score seciton pairs of bill dyad,
\textit{No. Alignments}: Number of alignments for bill dyad.} 
\label{tab:ideology_regs}
\end{table}